{\fontfamily{ptm}\selectfont Nesterov Type Momentum Methods}

### 2024-06-03 Notebook LaTEX/Momentum Accelerations/main.tex
new stuff

% ==============================================================================
\section{Preliminaries}
    In this section we list fundational results that are important for proofs in coming sections. 
    For this section, let the ambient space be $\RR^n$ and $\Vert \cdot\Vert$ be the Euclidean 2 norm until it's specified in the context. 
    \subsection{Lipschitz Smoothness}
        \begin{definition}[Lipschitz Smooth]
            Let $f$ be differentiable. 
            It has Lipschitz smoothness with constant $L$ if for all $x, y$
            $$
                \Vert \nabla f(x) - \nabla f(y)\Vert 
                \le 
                L \Vert x - y\Vert. 
            $$    
        \end{definition}

@@

    \subsection{Proximal Descent Inequality}
        \begin{theorem}[Proximal Descent Inequality]\label{thm:ppm_descent_ineq}
            With $f: \RR^n \mapsto \overline \RR^n$ $\beta$-convex where $\beta \ge 0$, fix any $x \in \RR^n$, let $p = \hprox_f(x)$, then for all $y$ we have inequality 
            $$
                \left(f(p) + \frac{1}{2}\Vert x - p\Vert^2\right)
                - 
                \left(
                    f(y) + \frac{1}{2}\Vert x - y\Vert^2 
                \right)
                \le 
                - \frac{(1 + \beta)}{2}\Vert y - p\Vert^2. 
            $$
            Recall $\hprox_\alpha f(x) = \argmin_{u}\left\lbrace f(u) + \frac{1}{2}\Vert u - x\Vert^2 \right\rbrace$. 
        \end{theorem}
        We make use of this theorem in the proof of convergence of proximal point method. 
        \begin{remark}
            This descent inequality can be generalized to bregman proximal mapping as well. 
        \end{remark}

\section{The Proximal Point Method in the Convex Case}
    In this section we go over the analysis of Proximal point method (PPM) in the convex case and see how the theories can be generalized into the cases where PPM is approximated. 
    \par
    With $f: \RR^n \mapsto \overline \RR$ lsc proper and convex, 
    given any $x_0$ the PPM generates sequence $(x_n)_{n\in \NN}$ by $x_{k + 1} = \hprox_{\eta_{k + 1}}f(x_k)$ for all $k \in \NN$ where the sequence $(\eta_{k})_{k \in \NN}$ is a nonegative sequence of real numbers.

@@

        \begin{theorem}[Convergence Rate of PPM (Strongly Convex)]
\section{Simple of the Proximal Point Method}

@@

    \subsection*{Generic Gradient Descent}

@@

            With $f$ be a function that has minimizer: $x_*$; $l_f(x; x_t)$ be a convex, lsc, proper lower-bounding function, define lower-bounding $\phi_t = \eta_{t + 1}l_f(x; x_t)$. 

@@

@@

            Fix any $x_0$, assume that the iterates $x_t$ defined for $t\in \NN$ satisfies

@@

            then the iterates satisfies: 

@@

            And for all $\eta_{t} \in (\epsilon, 2L^{-1} - \epsilon)$ with $\epsilon > 0$, the algorithm has sublinear convergence rates of

@@

                \frac{L - \epsilon^{-1}}{TL\epsilon} &\le 

@@

@@

@@

            For simplicity we make $\sigma_t = \sum_{i = 1}^{t}\eta_i$. 

@@

                \sigma_T (f(x_t) - f(y)) + \frac{1}{2}\Vert y - x_t\Vert^2

@@

                f(x_t) - f(y) + 

@@

                f(x_t) - f(y) &\le 

@@

            With $y = x_*$, we get the claimed convergence rate. 

@@

        \begin{remark}
            We speculate based on observing that the inequality 
            $$
                \phi_t(x) \le \eta_{t + 1}f(x) \le \phi_t(x) 
                + 
                \frac{L\eta_{t + 1}}{2}\Vert x - x_t\Vert^2 \quad \forall x \in \RR^n, 
            $$
            was invoked with $x = x_{t + 1}$ for the PPM descent inequality, meaning that having $\forall x \in \RR^n$ can be relaxed into $\forall (x_t)_{t \in \NN}$ generated by the algorithm, $\exists (L_t)_{t \in \NN}$, 
            $$
                \phi_t(x) \le \eta_{t + 1}f(x) \le \phi_t(x) 
                + 
                \frac{L_t\eta_{t + 1}}{2}\Vert x - x_t\Vert^2,
            $$
            where the sequence is generated by the algorithm. 
            This can be achieved by choosing the function $\Phi_{t + 1}$ at each iteration smartly. 
            \par
            The convergence rate is loose and when function $f$ exhibits additional favorable properties, such as being strongly convex, the convergence rate can be faster. 
        \end{remark}
    \subsection{Examples}
        \begin{example}[Proximal Gradient]
            In this section, we illustrate algorithms that satisfies the lower and uppwer bound estimate used in the above proof. 
            Consider $f = g + h$ with $h$ nonsmooth convex, and $g$ being $L$-Lipschitz smooth convex and differentiable. 
            Define $D_g(x, y) = g(x) - g(y) - \langle \nabla f(x), y - x\rangle$, which is the Bregman divergence of the function $g$. 
            Consider for all $x$: 
            \begin{align*}
                0 & \le 
                D_g (x, y) \le  \frac{L}{2} \Vert x - y\Vert^2 
                \\
                l_g (x; y) &\le 
                g(x) 
                \le l_g(x; y) + \frac{L}{2} \Vert x - y\Vert^2 
                \\
                h(x) + l_g(x; y) &\le f(x) = g(x) + h(x)
                \le 
                l_g(x; y) + h(x) + \frac{L}{2}\Vert x - y\Vert^2. 
            \end{align*}
            Define $\phi_t (x) = \eta_{t + 1}(h(x) + l_g(x; x_t))$, then results from previous theorems apply.    
        \end{example}
        \begin{remark}
            The envelope interpretation restricts the use of the theorem, since it requires that the proixmal operator is applied to the gradient of a function. 
            Extending the usage of the PPM descent inequality to other context requires operator theories and creativities. 
        \end{remark}

\section{Accelerated gradient descent}
    By recent works from (CITATION NEEDED), Nesterov accelerated gradient (CITATION NEEDED) can be interpreted as an approximation to the PPM method. 
    Additionally, works on accelerating PPM had bey done by Guler (CITATION NEEDED) decades ago. 
    The recent interpretations of Nesterov acceleration method via PPM focuses on the idea of a similar triangle, and unifying all varieties of Nesterov acceleration. 
    \subsection{The varieties of Nesterov Accelerated gradient}
        In this section, we list different varities of Nesterov accelerated method. 
        Each of the varieties will be presented generically because we are only interested in the forms they take.

    \subsection{Interpreting Nesterov accelerated gradient via PPM. }

>>>>

% ==============================================================================
\section{Preliminaries}
    In this section we list fundational results that are important for proofs in coming sections. 
    For this section, let the ambient space be $\RR^n$ and $\Vert \cdot\Vert$ be the Euclidean 2 norm until it's specified in the context. 
    \subsection{Lipschitz smoothness}
        \begin{definition}[Lipschitz Smooth]
            Let $f$ be differentiable. 
            It has Lipschitz smoothness with constant $L$ if for all $x, y$
            $$
                \Vert \nabla f(x) - \nabla f(y)\Vert 
                \le 
                L \Vert x - y\Vert. 
            $$    
        \end{definition}

@@

    \subsection{Proximal descent inequality}
        \begin{theorem}[Proximal Descent Inequality]\label{thm:ppm_descent_ineq}
            With $f: \RR^n \mapsto \overline \RR^n$ $\beta$-convex where $\beta \ge 0$, fix any $x \in \RR^n$, let $p = \hprox_f(x)$, then for all $y$ we have inequality 
            $$
                \left(f(p) + \frac{1}{2}\Vert x - p\Vert^2\right)
                - 
                \left(
                    f(y) + \frac{1}{2}\Vert x - y\Vert^2 
                \right)
                \le 
                - \frac{(1 + \beta)}{2}\Vert y - p\Vert^2. 
            $$
            Recall $\hprox_\alpha f(x) = \argmin_{u}\left\lbrace f(u) + \frac{1}{2}\Vert u - x\Vert^2 \right\rbrace$. 
        \end{theorem}
        We make use of this theorem in the proof of convergence of proximal point method. 
        \begin{remark}
            This descent inequality can be generalized to bregman proximal mapping as well. 
        \end{remark}

\section{The Proximal Point Method in the Convex Case}
    In this section we go over the analysis of Proximal point method (PPM) in the convex case and see how the theories can be generalized into the cases where PPM is approximated. 
    \subsection{Literature reviews}

    \par
    With $f: \RR^n \mapsto \overline \RR$ lsc proper and convex, 
    given any $x_0$ the PPM generates sequence $(x_n)_{n\in \NN}$ by $x_{k + 1} = \hprox_{\eta_{k + 1}}f(x_k)$ for all $k \in \NN$ where the sequence $(\eta_{k})_{k \in \NN}$ is a nonegative sequence of real numbers.

@@

\section{Applying the analysis of PPM}

@@

    \subsection{Generic gradient descent}

@@

            With $f$ convex having minimizer: $x_*$; $l_f(\cdot; x_t)$ convex, lsc and proper, define $\phi_t = \eta_{t + 1}l_f(x; x_t)$. 

@@

            \frac{L\eta_{t + 1}}{2}\Vert x - x_t\Vert^2 \quad \forall x \in \RR^n. 

@@

            Fix any $x_0$, let the iterates $x_t$ defined for $t\in \NN$ satisfies

@@

            then it has 

@@

            Additionally if $\exists \epsilon > 0: \eta_{t} \in (\epsilon, 2L^{-1} - \epsilon)$, for all $t \in \NN$, the algorithm has sublinear convergence rates of

@@

                f(x_T) - f(x_*)

@@

                \frac{L - \epsilon^{-1}}{TL\epsilon} (f(x_0) - f(x_*)) 

@@

            \par\noindent

@@

            For simplicity we set $\sigma_t := \sum_{i = 1}^{t}\eta_i$. 

@@

                \sigma_T (f(x_T) - f(y)) + \frac{1}{2}\Vert y - x_t\Vert^2

@@

                f(x_T) - f(y) + 

@@

                f(x_T) - f(y) &\le 

@@

            With $y = x_*$, we get the claimed convergence rate because $f(x_t)$ is strictly monotone decreasing. 

@@

            Observe that inequality 
            was invoked with $x = x_{t + 1}$ for the PPM descent inequality in the above proof, meaning that if  $\forall (x_t)_{t \in \NN}$ generated by the algorithm, $\exists (L_t)_{t \in \NN}$ such that
                \frac{L_t\eta_{t + 1}}{2}\Vert x - x_t\Vert^2,
            where the sequence is generated by the algorithm. 
            This can be achieved by choosing the function $\phi_t{t + 1}$ at each iteration smartly, then it's possible to still have the same convergenace rate. 
            In a practical setting, when $L_t = L$, and $\phi_{t + 1}(x) = \eta_{t + 1}f$, this is called a line search.
            The convergence rate is loose and when function $f$ exhibits additional favorable properties, such as being strongly convex, the convergence rate can be faster. 
    \subsection{Examples}
        \begin{example}[Proximal Gradient]
            In this section, we illustrate algorithms that satisfies the lower and uppwer bound estimate used in the above proof. 
            Consider $f = g + h$ with $h$ nonsmooth convex, and $g$ being $L$-Lipschitz smooth convex and differentiable. 
            Define $D_g(x, y) = g(x) - g(y) - \langle \nabla f(x), y - x\rangle$, which is the Bregman divergence of the function $g$. 
            Consider for all $x$: 
                0 & \le 
                D_g (x, y) \le  \frac{L}{2} \Vert x - y\Vert^2 
                l_g (x; y) &\le 
                g(x) 
                \le l_g(x; y) + \frac{L}{2} \Vert x - y\Vert^2 
                h(x) + l_g(x; y) &\le f(x) = g(x) + h(x)
                l_g(x; y) + h(x) + \frac{L}{2}\Vert x - y\Vert^2. 
            Define $\phi_t (x) = \eta_{t + 1}(h(x) + l_g(x; x_t))$, then results from previous theorems apply.    
        \end{example}
            The envelope interpretation restricts the use of the theorem, since it requires that the proixmal operator is applied to the gradient of a function. 
            Extending the usage of the PPM descent inequality to other context requires operator theories and creativities. 

\section{Accelerated gradient descent}
    By recent works from (CITATION NEEDED), Nesterov accelerated gradient (CITATION NEEDED) can be interpreted as an approximation to the PPM method. 
    Additionally, works on accelerating PPM had bey done by Guler (CITATION NEEDED) decades ago. 
    The recent interpretations of Nesterov acceleration method via PPM focuses on the idea of a similar triangle, and unifying all varieties of Nesterov acceleration. 
    \subsection{The varieties of Nesterov accelerated gradient}
        In this section, we list different varities of Nesterov accelerated method. 
        Each of the varieties will be presented generically because we are only interested in the forms they take.

    \subsection{Interpreting Nesterov accelerated gradient via PPM}
